### 2016-03-10 paper/icfp16-submission/semantics.tex
no induction in lemma k-stuck

\begin{lem}
\label{lem:k-stuck}
For all $t$,
if $\steptr{\trace}{\eapp{f}{\vhole{\thole}}}{\emptysu}{\emptysu}{e}{\vsu}{\tsu}$ and
   $\tincompat{t}{\ptype{\trace}{f}}$,
   then there exists a $v$ such that $\hastype{v}{t}$ and
   $\steps{\eapp{f}{v}}{\emptysu}{\emptysu}{\stuck}{\vsu'}{\tsu'}$ in at most
   $k$ steps, where $k$ is the length of $\trace$.
\end{lem}
\begin{proof}
  We can construct $v$ from $\trace$ by induction on $k$, the length of $\trace$.
  %
  The goal is to find the step at which the refinement of $\vhole{\thole}$,
  $\resolve{\vhole{\thole}}{\vsu}$ becomes incompatible with $t$.
  %
  The base case, $k = 0$ is trivial, $\vsu_0 = \emptysu$ and
  $\resolve{\vhole{\thole}}{\emptysu} = \vhole{\thole}$ which is compatible with
  \emph{any} type $t$.
  %
  In the inductive case, let
  $$\trace_{k-1} = \triple{\eapp{f}{\vhole{\thole}}}{\emptysu}{\emptysu},\ldots,\triple{e_{k-1}}{\vsu_{k-1}}{\tsu_{k-1}}$$
  be the prefix of $\trace$ of length $k-1$.
  %
  Let us split cases on whether $\resolve{\vhole{\thole}}{\vsu_{k-1}}$ is compatible with $t$.
  %
  \begin{description}
  \item[Case $\vincompat{\resolve{\vhole{\thole}}{\vsu_{k-1}}}{t}$:]
    %
    The inductive hypothesis applies.
    %
  \item[Case $\vcompat{\resolve{\vhole{\thole}}{\vsu_{k-1}}}{t}$
        but  $\vincompat{\resolve{\vhole{\thole}}{\vsu_{k}}}{t}$:]
    %
    In this case we will show that $\resolve{\vhole{\thole}}{\vsu_{k-1}}$
    must contain some other hole $\ehole'$ which is instantiated with a
    concrete value at step $k$.
    %
    Furthermore, $\ehole'$ is instantiated in such a way that
    $\vincompat{\resolve{\vhole{\thole}}{\vsu_{k}}}{t}$.
    %
    Finally, we will show that if we had instantiated $\ehole'$ with a
    value such that
    $\vcompat{\resolve{\vhole{\thole}}{\vsu_{k}}}{t}$, the current
    step would have gotten $\stuck$.

    % By Lemma~\ref{lem:fixme} we know that
    % $\vcompat{\resolve{\vhole{\thole}}{\vsu}}{\resolve{\thole}{\tsu}}$.
    %
    By Lemma~\ref{lem:vsu-ext} we know that
    $\vsu_{k} = \vsu_{k-1}[\ehole_1 \mapsto v_1] \ldots [\ehole_n \mapsto v_n]$.
    %
    We will assume, without loss of generality, that
    $\vsu_{k} = \vsu_{k-1}[\ehole' \mapsto v']$.

    Next, we introduce a notion of \emph{value contexts}, defined
    analogously to evaluation contexts.
    %
    $$
      V ::= \bullet \spmid \vnode{t}{V}{v}{v} \spmid \vnode{t}{v}{V}{v} \spmid \vnode{t}{v}{v}{V}
    $$
    %
    Since $\vsu_{k-1}$ and $\vsu_{k}$ differ only in $\ehole'$ but the resolved
    values differ,
    $\resolve{\vhole{\thole}}{\vsu_{k-1}} = \vcompat{\invctx{\ehole'[\thole']}}{t}$
    and
    $\resolve{\vhole{\thole}}{\vsu_{k}} = \vincompat{\invctx{v'}}{t}$.
    %
    Now, consider any value $u$ such that $\vcompat{\invctx{u}}{t}$,
    we show by case analysis on the evaluation relation that
    $\invctx{u}$ could not make progress at this step, and thus
    $\steps{\eapp{f}{\invctx{u}}}{\emptysu}{\emptysu}{\stuck}{\vsu}{\tsu}$.
  \end{description}
\end{proof}

>>>>

\begin{lem}
\label{lem:k-stuck}
For all $t$,
if $\steptr{\trace}{\eapp{f}{\vhole{\thole}}}{\emptysu}{\emptysu}{e}{\vsu}{\tsu}$ and
   $\tincompat{t}{\ptype{\trace}{f}}$,
   then there exists a $v$ such that $\hastype{v}{t}$ and
   $\steps{\eapp{f}{v}}{\emptysu}{\emptysu}{\stuck}{\vsu'}{\tsu'}$ in at most
   $k$ steps, where $k$ is the length of $\trace$.
\end{lem}
\begin{proof}
  We can construct $v$ from $\trace$ as follows.
  %
  Let
  $$
  \trace_i = \triple{\eapp{f}{\vhole{\thole}}}{\emptysu}{\emptysu},
             \ldots,
             \triple{e_{i-1}}{\vsu_{i-1}}{\tsu_{i-1}},
             \triple{e_{i}}{\vsu_{i}}{\tsu_{i}}
  $$
  be the shortest prefix of $\trace$ such that
  $\tincompat{\ptype{\trace_i}{f}}{t}$.
  %
  We will show that $\resolve{\vhole{\thole}}{\vsu_{i-1}}$
  must contain some other hole $\ehole'$ which is instantiated with a
  concrete value at step $i$.
  %
  Furthermore, $\ehole'$ is instantiated in such a way that
  $\vincompat{\resolve{\vhole{\thole}}{\vsu_{i}}}{t}$.
  \ES{need to connect \ptype{\trace{i}}{f} and \resolve{\vhole{\thole}}{\vsu_i}}
  %
  Finally, we will show that if we had instantiated $\ehole'$ with a
  value such that
  $\vcompat{\resolve{\vhole{\thole}}{\vsu_{i}}}{t}$, the current
  step would have gotten $\stuck$.

  % By Lemma~\ref{lem:fixme} we know that
  % $\vcompat{\resolve{\vhole{\thole}}{\vsu}}{\resolve{\thole}{\tsu}}$.
  %
  By Lemma~\ref{lem:vsu-ext} we know that
  $\vsu_{i} = \vsu_{i-1}[\ehole_1 \mapsto v_1] \ldots [\ehole_n \mapsto v_n]$.
  %
  We will assume, without loss of generality, that
  $\vsu_{i} = \vsu_{i-1}[\ehole' \mapsto v']$.

  Next, we introduce a notion of \emph{value contexts}, defined
  analogously to evaluation contexts.
  %
  $$
    V ::= \bullet \spmid \vnode{t}{V}{v}{v} \spmid \vnode{t}{v}{V}{v} \spmid \vnode{t}{v}{v}{V}
  $$
  %
  Since $\vsu_{i-1}$ and $\vsu_{i}$ differ only in $\ehole'$ but the resolved
  values differ,
  $\resolve{\vhole{\thole}}{\vsu_{i-1}} = \vcompat{\invctx{\ehole'[\thole']}}{t}$
  and
  $\resolve{\vhole{\thole}}{\vsu_{i}} = \vincompat{\invctx{v'}}{t}$.
  %
  Now, consider any value $u$ such that $\vcompat{\invctx{u}}{t}$,
  we show by case analysis on the evaluation relation that
  $\invctx{u}$ could not make progress at this step, and thus
  $\steps{\eapp{f}{\invctx{u}}}{\emptysu}{\emptysu}{\stuck}{\vsu'}{\tsu'}$.
\end{proof}
